---
title: "hw3"
author: "Lsy"
date: "2020/10/4"
output:
  word_document: default
  html_document: default
---

```{r,setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(norm)
library(mice)
library(metafor)
library(xlsx)
#set function
func_Hd <- function (CM, EM, CS, ES, CN, EN) { 
  sp <- sqrt ((((CN - 1) * CS ^ 2) + ((EN - 1) * ES ^ 2)) / (CN + EN - 2)) 
  d <- (EM - CM) / sp 
  Hd <- d * (1 - (3 / (4 * (CN + EN - 2) - 1))) 
  return (Hd) 
}
func_vHd <- function (Hd, CN, EN) {
  vHd <- ((CN + EN) / (CN * EN)) + (Hd ^ 2 / (2 * (CN + EN - 2))) 
  return (vHd)
}

setwd("D:/ELsy/2020fall/EDA/3/Lab3instruction")
hwData1 <- read.xlsx ("jpe0051-1669-sd1.xlsx",sheetName = "meta-analysis78 articles")
hwData2 <- read.xlsx ("jpe0051-1669-sd2.xlsx",sheetName = "meta-data 78",header = FALSE,startRow = 2)
colnames(hwData2) <- hwData2[1,]
hwdata <- hwData2 [-1,]
```

```{r,echo=FALSE, warning=FALSE}
# clear NA & NaN in variance and effect size
names(hwdata)[21] <-"efctsize"
names(hwdata)[23] <-"var"
chwdata <-subset(hwdata,!(hwdata["var"]=="NaN"))
chwdata1 <-subset(chwdata,!(chwdata["efctsize"]=="NA"))
hwdata <- subset(chwdata1,!(chwdata1["var"]=="NA"))
hwdata[, "unvar"]<- as.numeric(hwdata[,"var"])
hwdata[, "unefcts"]<- as.numeric(hwdata[,"efctsize"])
hwdata[, "tsh"]<- as.numeric(hwdata[,"post.years"])
hwdata[, "pro.ret"]<- as.numeric(hwdata[,"percent.retention"])
hwdata[, "unitsz"]<- as.numeric(hwdata[,"size"])
```



```{r,warning=FALSE}
# calculate mean effect size for different species richness and abundance
# for amphibians
amphibians <- subset(hwdata,taxon=="amphibians")
amphibiansR <- subset(amphibians,biodiversity=="richness")
ModelamphibianR <- rma(yi = unefcts, vi = unvar, method = "REML", data = amphibiansR)
amphibiansA <- subset(amphibians,biodiversity=="abundance")
ModelamphibianA <- rma(yi = unefcts, vi = unvar, method = "REML", data = amphibiansA)
summary(ModelamphibianR)
summary(ModelamphibianA)
```
#not sure why the I^2 here is 0.00?(So as to "beetles" etc.)

```{R}
# for birds
birds <- subset(hwdata,taxon=="birds")
birdsR <- subset(birds,biodiversity=="richness")
ModelbirdR <- rma(yi = unefcts, vi = unvar, method = "REML", data = birdsR)
birdsA <- subset(birds,biodiversity=="abundance")
ModelbirdA <- rma(yi = unefcts, vi = unvar, method = "REML", data = birdsA)
summary(ModelbirdR)
summary(ModelbirdA)
```
```{R,warning=FALSE}
# for beetles
beetles <- subset(hwdata,taxon=="beetles")
beetlesR <- subset(beetles,biodiversity=="richness")
ModelbeetleR <- rma(yi = unefcts, vi = unvar, method = "REML", data = beetlesR)
beetlesA <- subset(beetles,biodiversity=="abundance")
ModelbeetleA <- rma(yi = unefcts, vi = unvar, method = "REML", data = beetlesA)
summary(ModelbeetleR)
summary(ModelbeetleA)
```
```{R,warning=FALSE}
# for bryophytes
bryophytes <- subset(hwdata,taxon=="bryophytes")
bryophytesR <- subset(bryophytes,biodiversity=="richness")
ModelbryophyteR <- rma(yi = unefcts, vi = unvar, method = "REML", data = bryophytesR)
bryophytesA <- subset(bryophytes,biodiversity=="abundance")
ModelbryophyteA <- rma(yi = unefcts, vi = unvar, method = "REML", data = bryophytesA)
summary(ModelbryophyteR)
summary(ModelbryophyteA)
```
```{R,warning=FALSE}
# for fungi
fungi <- subset(hwdata,taxon=="fungi")
fungiR <- subset(fungi,biodiversity=="richness")
ModelfungiR <- rma(yi = unefcts, vi = unvar, method = "REML", data = fungiR)
fungiA <- subset(fungi,biodiversity=="abundance")
ModelfungiA <- rma(yi = unefcts, vi = unvar, method = "REML", data = fungiA)
summary(ModelfungiR)
summary(ModelfungiA)
```
```{R,warning=FALSE}
# for invertebrates
invertebrates <- subset(hwdata,taxon=="invertebrates")
invertebratesR <- subset(invertebrates,biodiversity=="richness")
ModelinvertebratesR <- rma(yi = unefcts, vi = unvar, method = "REML", data = invertebratesR)
invertebratesA <- subset(invertebrates,biodiversity=="abundance")
ModelinvertebratesA <- rma(yi = unefcts, vi = unvar, method = "REML", data = invertebratesA)
summary(ModelinvertebratesR)
summary(ModelinvertebratesA)
```
```{R,warning=FALSE}
# for lichens
lichens <- subset(hwdata,taxon=="lichens")
lichensR <- subset(lichens,biodiversity=="richness")
ModellichenR <- rma(yi = unefcts, vi = unvar, method = "REML", data = lichensR)
lichensA <- subset(lichens,biodiversity=="abundance")
ModellichenA <- rma(yi = unefcts, vi = unvar, method = "REML", data = lichensA)
summary(ModellichenR)
summary(ModellichenA)
```
```{R,warning=FALSE}
# for mammals
mammals <- subset(hwdata,taxon=="mammals")
mammalsR <- subset(mammals,biodiversity=="richness")
ModelmammalR <- rma(yi = unefcts, vi = unvar, method = "REML", data = mammalsR)
mammalsA <- subset(mammals,biodiversity=="abundance")
ModelmammalA <- rma(yi = unefcts, vi = unvar, method = "REML", data = mammalsA)
summary(ModelmammalR)
summary(ModelmammalA)
```
```{R}
# for plants
plants <- subset(hwdata,taxon=="plants")
plantsR <- subset(plants,biodiversity=="richness")
ModelplantR <- rma(yi = unefcts, vi = unvar, method = "REML", data = plantsR)
plantsA <- subset(plants,biodiversity=="abundance")
ModelplantA <- rma(yi = unefcts, vi = unvar, method = "REML", data = plantsA)
summary(ModelplantR)
summary(ModelplantA)
```

```{R}
# for spiders
spiders <- subset(hwdata,taxon=="spiders")
spidersR <- subset(spiders,biodiversity=="richness")
ModelspiderR <- rma(yi = unefcts, vi = unvar, method = "REML", data = spidersR)
spidersA <- subset(spiders,biodiversity=="abundance")
ModelspiderA <- rma(yi = unefcts, vi = unvar, method = "REML", data = spidersA)
summary(ModelspiderR)
summary(ModelspiderA)
```

```{R}
# for reptiles
reptiles <- subset(hwdata,taxon=="reptiles")
reptilesR <- subset(reptiles,biodiversity=="richness")
ModelreptileR <- rma(yi = unefcts, vi = unvar, method = "REML", data = reptilesR)
reptilesA <- subset(reptiles,biodiversity=="abundance")
ModelreptileA <- rma(yi = unefcts, vi = unvar, method = "REML", data = reptilesA)
summary(ModelreptileR)
summary(ModelreptileA)
```
```{R,warning=FALSE}
# meta-regression for examine influence of moderators
#for forest biome
Modelbiome1 <- rma (yi = unefcts, vi = unvar, mods = ~ temperate.boreal, method = "REML", data = hwdata) 
summary (Modelbiome1)
Modelbiome2 <- rma (yi = unefcts, vi = unvar, mods = ~ temperate.boreal-1, method = "REML", data = hwdata) 
summary (Modelbiome2)
```
```{R,warning=FALSE}
#for time since harvest
Modelposty <- rma (yi = unefcts, vi = unvar, mods = ~ tsh, method = "REML", data = hwdata) 
summary (Modelposty)
Modelposty2 <- rma (yi = unefcts, vi = unvar, mods = ~ tsh -1, method = "REML", data = hwdata) 
summary (Modelposty2)
```

```{R,warning=FALSE}
#for proportion of retention
Modelpro.ret <- rma (yi = unefcts, vi = unvar, mods = ~ pro.ret, method = "REML", data = hwdata) 
summary (Modelpro.ret)
Modelpro.ret2 <- rma (yi = unefcts, vi = unvar, mods = ~ pro.ret -1, method = "REML", data = hwdata) 
summary (Modelpro.ret2)

#mix all the variables
Modelmix <- rma (yi = unefcts, vi = unvar, mods = ~ pro.ret + unitsz + tsh + temperate.boreal, method = "REML", data = hwdata) 
summary (Modelmix)
Modelmix2 <- rma (yi = unefcts, vi = unvar, mods = ~ pro.ret + unitsz + tsh + temperate.boreal - 1, method = "REML", data = hwdata) 
summary (Modelmix2)
```
The effect size is sinificantly influenced by unit size and forest biome (samples from boreal biome estimate larger than that from temperate biome).
```{R,warning=FALSE}
#test publication bias
regtest(Modelmix,model = "lm")
ranktest(Modelmix)
```
There is not significant asymetry.

According to the results above, the effect size is mostly influenced by unit size and forest biome (samples from boreal biome estimate larger than that from temperate biome).