---
title: "HW10"
author: "Lsy-2020215195"
date: "2020/11/21"
output: word_document
---

```{r setup, include=FALSE}
setwd("D:/ELsy/2020fall/EDA/10/Labnotes/extdata")
library("dplyr")
library("tidyr")
library("STRbook")
```

```{r}
#data processing
locs <- read.table(system.file("extdata", "Stationinfo.dat",
                               package = "STRbook"),
                  col.names = c("id", "lat", "lon"))
Times <- read.table(system.file("extdata", "Times_1990.dat",
                                package = "STRbook"),
                  col.names = c("julian", "year", "month", "day"))
Tmax <- read.table(system.file("extdata", "Tmax_1990.dat",
                               package = "STRbook"))
names(Tmax) <- locs$id
Tmax <- cbind(Times, Tmax)
head(names(Tmax), 10)
Tmax_long <- gather(Tmax, id, z, -julian, -year, -month, -day)
head(Tmax_long)
Tmax_long$id <- as.integer(Tmax_long$id)
nrow(Tmax_long)
Tmax_long <- filter(Tmax_long, !(z <= -9998))
nrow(Tmax_long)
Tmax_long <- mutate(Tmax_long, proc = "Tmax")
head(Tmax_long)

data(Tmin_long, package = "STRbook")
data(TDP_long, package = "STRbook")
data(Precip_long, package = "STRbook")
NOAA_df_1990 <- rbind(Tmax_long, Tmin_long, TDP_long, Precip_long)

set.seed(1)
data("NOAA_df_1990", package = "STRbook")
Tmax <- filter(NOAA_df_1990,     # subset the data
              proc == "Tmax" &   # only max temperature
              month %in% 5:9 &   # May to September
              year == 1993)      # year of 1993

Tmax %>% select(lon, lat, date, julian, z) %>% head()
Tmax$t <- Tmax$julian - 728049     # create a new time variable

```
## 1. Hovmöller plot based on longititude and time
```{r,eval=FALSE}
#generate grids
lim_lon <- range(Tmax$lon)        # longitude range
lim_t <- range(Tmax$t)            # time range
lon_axis <- seq(lim_lon[1],       # longitude axis
                lim_lon[2],
                length=20)
t_axis <- seq(lim_t[1],           # time axis
              lim_t[2],
              length=100)
lon_t_grid <- expand.grid(lon = lon_axis,
                          t = t_axis)
Tmax_grid_lon <- Tmax
dist_lon <- abs(outer(Tmax$lon, lon_axis, "-"))
Tmax_grid_lon$lon <- lon_axis[apply(dist_lon, 1, which.min)]

Tmax_lon_Hov <- group_by(Tmax_grid_lon, lon, t) %>%
                summarise(z = mean(z))

Hovmoller_lon <- ggplot(Tmax_lon_Hov) +            
        geom_tile(aes(x = lon, y = t, fill = z)) +
        fill_scale(name = "degF") +     
        scale_y_reverse() +           
        ylab("Day number (days)") +     
        xlab("longitude (degrees)") +    
        theme_bw()                    
station <- 13966
station_lon <- filter(Tmax,id == 13966)$lon[1]
station_lat <- filter(Tmax,id == 13966)$lat[1]
Hovmoller_lon <- Hovmoller_lon + geom_vline(xintercept = station_lon,linetype='dashed')
plot(Hovmoller_lon)
```
According to the Hovmöller plot, the trend of data during the observing period is similar.

## 2. Plot predict result from section 2.3.
```{r,eval=FALSE}
# continue with Lecture10 demo labnotes
data("SOI", package = "STRbook")
SOI_df <- select(SOI, -Ann) %>% gather(Month, soi, -Year)
soi_pred <- filter(SOI_df, Month == "Oct" & Year == "1997") %>%
            select(soi)

predict_one_pixel <- function(lm, soi_pred) {
    predict(lm,                           
            newdata = soi_pred,          
            interval = "prediction") %>%  
    data.frame() %>%                      
    mutate(se = (upr-lwr)/(2 * 1.96)) %>% 
    select(fit, se)                       
  }

SST_Oct_1997 <- pixel_lms %>%
                mutate(preds = map(model,
                                   predict_one_pixel,
                                   soi_pred = soi_pred)) %>%
                unnest(preds)
SST_Oct_1997 %>% head(3) 

################
#Plor predict result
predictSST_plot <- ggplot(SST_Oct_1997) +
    geom_tile(aes(x = lon,y = lat,       
                  fill=fit)) +
    fill_scale() + coord_fixed() + theme_bw()
print(predictSST_plot)
```
As the plotted prediction shows, there might be a significant rise of SST in the east pacific in Oct 1997(which is also congruent with the fact).


## 3. Refit the GLM model in section 2.4
```{R}
#here we use quasi-Poisson account for over-dispersion
#read and fit data
data("MOcarolinawren_long", package = "STRbook")
MOcarolinawren_long <- MOcarolinawren_long %>%
                       filter(!is.na(cnt))
G <- auto_basis(data = MOcarolinawren_long[,c("lon","lat")] %>%
                       SpatialPoints(),          
                nres = 1,                         
                type = "Gaussian")                

S <- eval_basis(basis = G,                      
                s = MOcarolinawren_long[,c("lon","lat")] %>%
                     as.matrix()) %>%           
     as.matrix()                                
colnames(S) <- paste0("B", 1:ncol(S))

#data processing
#attach the basis-function 
Wren_df <- cbind(MOcarolinawren_long,S) %>%
  select(-loc.ID, -t)
Wren_df[1:3, 1:5]
#fit a GLM model (use quasipoisson for over-dispersion)
Wren_GLM_qP_qP <- glm(cnt ~ (lon + lat + year)^2 + ., 
                family = quasipoisson("log"),     
                data = Wren_df)             

options(digits = 3)

###########
#Prediction
pred_grid <- expand.grid(lon = seq(
                             min(MOcarolinawren_long$lon) - 0.2,
                             max(MOcarolinawren_long$lon) + 0.2,
                             length.out = 80),
                         lat = seq(
                             min(MOcarolinawren_long$lat) - 0.2,
                             max(MOcarolinawren_long$lat) + 0.2,
                             length.out = 80),
                         year = 1994:2014)

#evaluate basis functions
S_pred <- eval_basis(basis = G,                   
                s = pred_grid[,c("lon","lat")] %>% 
                     as.matrix()) %>%            
     as.matrix()                                 
colnames(S_pred) <- paste0("B", 1:ncol(S_pred))  
pred_grid <- cbind(pred_grid,S_pred)             

wren_preds <- predict(Wren_GLM_qP,
                      newdata = pred_grid,
                      type = "link",
                      se.fit = TRUE)
pred_grid <- pred_grid %>%
             mutate(log_cnt = wren_preds$fit,
                    se = wren_preds$se.fit)

#plot predictions & se
g1_qP <- ggplot() + geom_raster(data=pred_grid,
                             aes(lon, lat, fill = pmax(pmin(log_cnt,4),0))) +
  facet_wrap(~year,nrow=3,ncol=7) +
    geom_point(data = filter(MOcarolinawren_long, !is.na(cnt)),
               aes(lon, lat),colour="black", size=3) +
    geom_point(data=filter(MOcarolinawren_long,!is.na(cnt)),aes(lon,lat,colour=log(cnt)),size=2) +
    scale_colour_distiller(palette="Spectral",limits=c(0,4)) +
    scale_fill_distiller(palette="Spectral",limits=c(0,4),name=expression(log(Y[t]))) + theme_bw()
g1_qP

g2_qP <- ggplot() + geom_raster(data=pred_grid,aes(lon,lat,fill=se)) +
  facet_wrap(~year,nrow=3,ncol=7) +
    scale_fill_distiller(palette="BrBG",limits=c(0,1),name=expression(s.e.)) + theme_bw()
g2_qP

Wren_df$residuals <- residuals(Wren_GLM_qP)
g3_qP <- ggplot(Wren_df) +
    geom_point(aes(lon, lat, colour = residuals)) +
    col_scale(name = "residuals") +
    facet_wrap(~year, nrow = 3) + theme_bw()
g3_qP

P <- list()                                
years <- 1994:2014
for(i in seq_along(years)) {                
  Wren_year <- filter(Wren_df,
                     year == years[i])      
  obs_dists <- Wren_year %>%                
    select(lon,lat) %>%                     
    dist() %>%                              
    as.matrix()                            
  obs_dists.inv <- 1/obs_dists             
  diag(obs_dists.inv) <- 0                
  P[[i]] <- Moran.I(Wren_year$residuals,   
                    obs_dists.inv) %>%
            do.call("cbind", .)            
}
do.call("rbind",P) %>% summary(digits = 2)

```

Modle using quasiPoisson shows similar result as the original Poisson one, despite the higher coastal value in g2_qP plot(stands for prediction standard errors).
